--- 
title: "Civil Liberties and Inequalities"
author: "Léo Persoz"
format: 
  dashboard:
    css: styles.css
    logo: images/LOGO_IMAGE.png
    nav-buttons: [github]
    github: https://github.com/YOUR_URL
theme: lux
execute:
  echo: false
  warning: false
  message: false
---

``` {r load/install packages and datasets}
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, 
               here,
               sf,
               bslib, 
               bsicons,
               rnaturalearth, 
               plotly, 
               countrycode, 
               htmltools, 
               reactable,
               janitor,
               countrycode
               #shiny,
               #car,
               #mgcv
               )

gini_raw <- read_csv(here("data/gini_coef.csv.csv"))

gni_cap_raw <- read_csv(here("data/gnicap_atm_con.csv"))

civil_lib_raw <- read_csv(here("data/cliberties_idea.csv"))

#general option of the document
options(scipen = 999)
```

```{r select years and pivot dataframe}
gini_long <- gini_raw %>%
  select(1, 147:225) %>%
  pivot_longer(cols = 2:80,
               names_to = "year", 
               values_to = "value"
               ) %>% 
  rename(gini_index = value) %>% 
  mutate(gini_index = gini_index/100)

gni_cap_long <- gni_cap_raw %>%
  mutate(across(2:ncol(gni_cap_raw), ~str_replace_all(., "k", "e3"))) %>%
  mutate(across(2:ncol(gni_cap_raw), ~str_replace_all(., "M", "e6"))) %>%
  pivot_longer(cols = 2:252,
               names_to = "year",
               values_to = "value") %>%
  mutate(value = as.numeric(value)) %>% 
  rename(gni_cap = value)

civil_lib_long <- civil_lib_raw %>%
  pivot_longer(cols = 2:48,
               names_to = "year",
               values_to = "value") %>% 
  rename(civlib_index = value)
```

```{r}
# value box - most inequality and least civil liberties
top_highest_gini <- gini_long %>%
  filter(year == "2023") %>%
  mutate(country = case_when(country == "Central African Republic" ~ "Central Afr. Rep.", TRUE ~ country)) %>%
  arrange(desc(gini_index)) %>% 
  head(5)
  
top_highest_gini$country <- factor(top_highest_gini$country, levels = top_highest_gini$country)

sparkcol_highest_gini <- plot_ly(top_highest_gini) %>%
  add_bars(
    x = ~country, y = ~gini_index,
    text = ~country,
    textposition = "inside",
    textfont = list(color = "darkgrey"),
    insidetextanchor = "middle",
    color = I("white"), 
    alpha = 0.8
  ) %>%
  layout(
    xaxis = list(visible = F, showgrid = F, title = ""),
    yaxis = list(visible = T, showgrid = F, title = ""),
    hovermode = "x",
    margin = list(t = 0, r = 0, l = 0, b = 0),
    font = list(color = "white"),
    paper_bgcolor = "transparent",
    plot_bgcolor = "transparent"
  ) %>%
  config(displayModeBar = F) %>%
  htmlwidgets::onRender(
    "function(el) {
      el.closest('.bslib-value-box')
        .addEventListener('bslib.card', function(ev) {
          Plotly.relayout(el, {'xaxis.visible': ev.detail.fullScreen});
        })
    }"
  )

# checking how skewed the gini_index variable is
gini_index_hist <- gini_long %>%
  filter(year == "2023") %>%
  ggplot(aes(x = gini_index)) +
  geom_histogram(binwidth = 0.01, fill = "blue", color = "black") #sort of skewed but okay

# taking the mean (0.4) and not the median (0.38) because data isn't so skewed
mean_gini <- gini_long %>%
  filter(year == "2023") %>%
  summarize(mean_gini = round(mean(gini_index), 2))

#value_box(
  #title = "Top 5 country with the highest Gini Index, 2023",
  #value = "Highest Gini Index: South Africa (0.66)",
  #showcase = sparkcol_highest_gini,
  #full_screen = TRUE,
  #theme = value_box_theme(bg = "#1E90FF")
#)
```

```{r value box, highest/lowest civil liberties}
top_highest_civlib <- civil_lib_long %>%
  filter(year == "2021") %>%
  arrange(desc(civlib_index)) %>% 
  head(5)
  
top_highest_civlib$country <- factor(top_highest_civlib$country, levels = top_highest_civlib$country)

sparkcol_highest_civlib <- plot_ly(top_highest_civlib) %>%
  add_bars(
    x = ~country, y = ~civlib_index,
    text = ~country,
    textposition = "inside",
    textfont = list(color = "darkgrey"),
    insidetextanchor = "middle",
    color = I("white"), 
    alpha = 0.8
  ) %>%
  layout(
    xaxis = list(visible = F, showgrid = F, title = ""),
    yaxis = list(visible = T, showgrid = F, title = ""),
    hovermode = "x",
    margin = list(t = 0, r = 0, l = 0, b = 0),
    font = list(color = "white"),
    paper_bgcolor = "transparent",
    plot_bgcolor = "transparent"
  ) %>%
  config(displayModeBar = F) %>%
  htmlwidgets::onRender(
    "function(el) {
      el.closest('.bslib-value-box')
        .addEventListener('bslib.card', function(ev) {
          Plotly.relayout(el, {'xaxis.visible': ev.detail.fullScreen});
        })
    }"
  )

top_lowest_civlib <- civil_lib_long %>%
  filter(year == "2021") %>%
  arrange(civlib_index) %>% 
  head(5)
  
top_lowest_civlib$country <- factor(top_lowest_civlib$country, levels = top_lowest_civlib$country)

sparkcol_lowest_civlib <- plot_ly(top_lowest_civlib) %>%
  add_bars(
    x = ~country, y = ~civlib_index,
    text = ~country,
    textposition = "inside",
    textfont = list(color = "darkgrey"),
    insidetextanchor = "middle",
    color = I("white"), 
    alpha = 0.8
  ) %>%
  layout(
    xaxis = list(visible = F, showgrid = F, title = ""),
    yaxis = list(visible = T, showgrid = F, title = ""),
    hovermode = "x",
    margin = list(t = 0, r = 0, l = 0, b = 0),
    font = list(color = "white"),
    paper_bgcolor = "transparent",
    plot_bgcolor = "transparent"
  ) %>%
  config(displayModeBar = F) %>%
  htmlwidgets::onRender(
    "function(el) {
      el.closest('.bslib-value-box')
        .addEventListener('bslib.card', function(ev) {
          Plotly.relayout(el, {'xaxis.visible': ev.detail.fullScreen});
        })
    }"
  )

```


```{r worldmap + joining gini and map}
country_shapess <- rnaturalearth::ne_countries()

gini_long <- gini_long %>% 
  mutate(
    country_code = countrycode(country, 
                               "country.name", 
                               "iso3c"),
    continent = countrycode(country, 
                            "country.name",
                            "continent"))

# setdiff(gini_long$country_code, country_shapess$adm0_a3)
# setdiff(country_shapess$adm0_a3, gini_long$country_code) 

#joining gini and geometry
geom_x_gini <- left_join(country_shapess, gini_long, by = c("adm0_a3" = "country_code")) %>% 
  mutate(tooltip_label = paste(country,
                               round(gini_index, 2),
                               sep = ": ")) %>%
  filter(year == "2023")

#map fill gini 2023
geom_x_gini_map <- 
  geom_x_gini %>% 
  ggplot() +
  geom_sf(aes(fill = gini_index, text = tooltip_label, geometry = geometry)) +
  labs(title = "Gini index (GI) per country, 2023",
       fill = "GI") +
  coord_sf(expand = F)

map_gini_plotly <- ggplotly(geom_x_gini_map, tooltip = "text")
```
```{r map civil liberties}
country_shapess <- rnaturalearth::ne_countries()

civil_lib_long <- civil_lib_long %>% 
  mutate(
    country_code = countrycode(country, 
                               "country.name", 
                               "iso3c"),
    continent = countrycode(country, 
                            "country.name",
                            "continent"))

# setdiff(civil_lib_long$country_code, country_shapess$adm0_a3)
# setdiff(country_shapess$adm0_a3, civil_lib_long$country_code) 

#joining civil liberty index and geometry
geom_x_civlib <- left_join(country_shapess, civil_lib_long, by = c("adm0_a3" = "country_code")) %>% 
  mutate(tooltip_label = paste(country,
                               civlib_index,
                               sep = ": "))

# map fill gini 2023
geom_x_civlib_map <- 
  geom_x_civlib %>% 
  ggplot() +
  geom_sf(aes(fill = civlib_index, text = tooltip_label, geometry = geometry)) +
  scale_fill_gradient(low = "lightgreen", high = "darkgreen") +
  labs(title = "Civil liberties Index (CLI) per country, 2021",
       fill = "CLI") +
  coord_sf(expand = F)

map_civlib_plotly <- ggplotly(geom_x_civlib_map, tooltip = "text")
```


```{r column graphs}
highest_gini_2023 <- gini_long %>% 
  filter(year == "2023") %>% 
  arrange(desc(gini_index)) %>% 
  head(20)

highest_gini_2023_plot <- highest_gini_2023 %>% 
  ggplot() +
  geom_col(aes(y = gini_index, x = reorder(country, gini_index), fill = gini_index)) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(title = "Countries with highest Gini index, 2023",
       y = "Country",
       x = "Gini Index")

highest_gini_2023_plotly <- ggplotly(highest_gini_2023_plot) %>%
  layout(xaxis = list(tickangle = 45))

highest_gini_2000 <- gini_long %>% 
  filter(year == "2000") %>% 
  arrange(desc(gini_index)) %>% 
  head(20)

highest_gini_2000_plot <- highest_gini_2000 %>% 
  ggplot() +
  geom_col(aes(y = gini_index, x = reorder(country, gini_index), fill = gini_index)) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(title = "Countries with highest Gini index, 2000",
       y = "Country",
       x = "Gini Index")

highest_gini_2000_plotly <- ggplotly(highest_gini_2000_plot) %>%
  layout(xaxis = list(tickangle = 45))
```


```{r line graph mean gini}
# world's mean
mean_gini_year <- gini_long %>% 
  group_by(year) %>% 
  summarise(mean_gini = mean(gini_index, na.rm = T))

mean_gini_year <- 
  mean_gini_year %>% 
  mutate(year = as.character(year)) %>%
  mutate(mean_gini = as.numeric(mean_gini))

# round(mean_gini_year$mean_gini, 3)

# continents means
mean_gini_year_asia <- gini_long %>% 
  filter(continent == "Asia") %>% 
  group_by(year) %>% 
  summarise(mean_gini_asia = mean(gini_index, na.rm = T))

mean_gini_year_africa <- gini_long %>% 
  filter(continent == "Africa") %>% 
  group_by(year) %>% 
  summarise(mean_gini_afr = mean(gini_index, na.rm = T))

mean_gini_year_EU <- gini_long %>% 
  filter(continent == "Europe") %>% 
  group_by(year) %>% 
  summarise(mean_gini_EU = mean(gini_index, na.rm = T))

mean_gini_year_ocea <- gini_long %>% 
  filter(continent == "Oceania") %>% 
  group_by(year) %>% 
  summarise(mean_gini_ocea = mean(gini_index, na.rm = T))

#creating north america and latin america (to avoid biases coming from economic differences between regions)
gini_long <- gini_long %>% 
  mutate(continent = case_when(country %in% c("USA", "Canada") & continent == "Americas" ~ "North America",
                               !(country %in% c("USA", "Canada")) & continent == "Americas" ~ "Latin America", T ~ continent))

mean_gini_year_northAM <- gini_long %>% 
  filter(continent == "North America") %>% 
  group_by(year) %>% 
  summarise(mean_gini_northam = mean(gini_index, na.rm = T))

mean_gini_year_latAM <- gini_long %>% 
  filter(continent == "Latin America") %>% 
  group_by(year) %>% 
  summarise(mean_gini_latam = mean(gini_index, na.rm = T))

# joining world's mean with regions' means in one data frame
mean_gini_year_all <- left_join(mean_gini_year, mean_gini_year_africa)
mean_gini_year_all <- left_join(mean_gini_year_all, mean_gini_year_asia)
mean_gini_year_all <- left_join(mean_gini_year_all, mean_gini_year_EU)
mean_gini_year_all <- left_join(mean_gini_year_all, mean_gini_year_northAM)
mean_gini_year_all <- left_join(mean_gini_year_all, mean_gini_year_latAM)
mean_gini_year_all <- left_join(mean_gini_year_all, mean_gini_year_ocea)

mean_gini_year_all <- mean_gini_year_all %>% 
  mutate(year = as.numeric(year))

# covert to long format, adapt the label of the tooltip for each variable for the ggplotly call
mean_gini_long <- mean_gini_year_all %>%
  pivot_longer(cols = starts_with("mean_gini"), names_to = "region", values_to = "mean_gini") %>%
  mutate(region = case_when(
    region == "mean_gini" ~ "World",
    region == "mean_gini_afr" ~ "Africa",
    region == "mean_gini_asia" ~ "Asia",
    region == "mean_gini_EU" ~ "Europe",
    region == "mean_gini_northam" ~ "North America",
    region == "mean_gini_latam" ~ "Latin America",
    region == "mean_gini_ocea" ~ "Oceania",
    TRUE ~ region),
  tooltip_label = paste(region, " (", year, "):", round(mean_gini, 2), sep = ""))

# creating the line graph
graph_mean_gini_all <- 
  mean_gini_long %>% 
  ggplot() +
  geom_line(aes(x = year, y = mean_gini, color = region, text = tooltip_label, group = region)) +
  scale_x_continuous(breaks = seq(min(mean_gini_year_all$year), max(mean_gini_year_all$year), by = 5),
                     guide = guide_axis(angle = 45)) +
  scale_y_continuous(limits = c(0.28, 0.5)) +
  labs(title = "Mean Gini index per year (world and regions)",
       x = "Year",
       y = "Mean Gini index",
       color = "Region") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7))

graph_mean_gini_all_plotly <- ggplotly(graph_mean_gini_all, tooltip = "text") %>%
  layout(xaxis = list(tickangle = 45))
```

```{r dataframe gini, civlib and GNI}
# adding country code and continent to the gni dataframe and filtering gini dataframe to 1975
gni_cap_long <- gni_cap_long %>%
  mutate(
    country_code = countrycode(country, 
                               "country.name", 
                               "iso3c"),
    continent = countrycode(country, 
                            "country.name",
                            "continent")) %>% 
  filter(between(as.numeric(year), 1975, 2021))

gini_long_75 <- gini_long %>% 
  filter(between(as.numeric(year), 1975, 2021))

# joining the gini, civil liberties and gni dataframe

#setdiff(gini_long_75$country_code, civil_lib_long$country_code)
#setdiff(civil_lib_long$country_code, gini_long_75$country_code)

intermed_join <- left_join(gini_long_75, civil_lib_long, by = c("country_code", "year"))

intermed_join <- intermed_join %>%
  select(-matches("\\.y")) %>% 
  rename(continent = continent.x,
         country = country.x)

gini_civlib_gni <- left_join(intermed_join, gni_cap_long, by = c("country_code", "year"))

gini_civlib_gni <- gini_civlib_gni %>%
  select(-matches("\\.y")) %>% 
  rename(continent = continent.x,
         country = country.x)

gini_civlib_gni <- gini_civlib_gni %>%
  mutate(year = as.numeric(year), 
         gini_index = as.numeric(gini_index),
         civlib_index = as.numeric(civlib_index),
         gni_cap = as.numeric(gni_cap))
```

```{r 3D graph: dataframe + data wrangling}
gini_civlib_gni_asia <- gini_civlib_gni %>% 
  filter(continent == "Asia") %>% 
  group_by(year) %>% 
  summarise(mean_gini_asi = mean(gini_index, na.rm = T),
            mean_civlib_asi = mean(civlib_index, na.rm = T),
            mean_gni_asi = mean(gni_cap, na.rm = T))

gini_civlib_gni_africa <- gini_civlib_gni %>% 
  filter(continent == "Africa") %>% 
  group_by(year) %>% 
  summarise(mean_gini_afr = mean(gini_index, na.rm = T),
            mean_civlib_afr = mean(civlib_index, na.rm = T),
            mean_gni_afr = mean(gni_cap, na.rm = T))

gini_civlib_gni_northam <- gini_civlib_gni %>% 
  filter(continent == "North America") %>% 
  group_by(year) %>% 
  summarise(mean_gini_northam = mean(gini_index, na.rm = T),
            mean_civlib_northam = mean(civlib_index, na.rm = T),
            mean_gni_northam = mean(gni_cap, na.rm = T))

gini_civlib_gni_latam <- gini_civlib_gni %>% 
  filter(continent == "Latin America") %>% 
  group_by(year) %>% 
  summarise(mean_gini_latam = mean(gini_index, na.rm = T),
            mean_civlib_latam = mean(civlib_index, na.rm = T),
            mean_gni_latam = mean(gni_cap, na.rm = T))

gini_civlib_gni_ocea <- gini_civlib_gni %>% 
  filter(continent == "Oceania") %>% 
  group_by(year) %>% 
  summarise(mean_gini_ocea = mean(gini_index, na.rm = T),
            mean_civlib_ocea = mean(civlib_index, na.rm = T),
            mean_gni_ocea = mean(gni_cap, na.rm = T))

gini_civlib_gni_eu <- gini_civlib_gni %>% 
  filter(continent == "Europe") %>% 
  group_by(year) %>% 
  summarise(mean_gini_eu = mean(gini_index, na.rm = T),
            mean_civlib_eu = mean(civlib_index, na.rm = T),
            mean_gni_eu = mean(gni_cap, na.rm = T))
  

 # joining datasets
mean_gini_civ_gni <- left_join(gini_civlib_gni_asia, gini_civlib_gni_africa)
mean_gini_civ_gni <- left_join(mean_gini_civ_gni, gini_civlib_gni_northam)
mean_gini_civ_gni <- left_join(mean_gini_civ_gni, gini_civlib_gni_latam)
mean_gini_civ_gni <- left_join(mean_gini_civ_gni, gini_civlib_gni_ocea)
mean_gini_civ_gni <- left_join(mean_gini_civ_gni, gini_civlib_gni_eu)

# pivoting and renaming variables
mean_gini_civ_gni_long <- mean_gini_civ_gni %>%
  pivot_longer(cols = starts_with("mean_"), 
               names_to = "names", 
               values_to = "value") %>% 
  separate(names, into = c("measure", "continent"), sep = "_", extra = "merge") %>% 
  separate(continent, into = c("measure_type", "continent"), sep = "_") %>% 
  select(-measure) %>% 
  pivot_wider(names_from = measure_type, values_from = value) %>% 
  mutate(continent = case_when(continent == "asi" ~ "Asia",
                               continent == "afr" ~ "Africa",
                               continent == "northam" ~ "N. America",
                               continent == "latam" ~ "L. America",
                               continent == "ocea" ~ "Oceania",
                               continent == "eu" ~ "Europe")) %>% 
  mutate(gini = round(gini, 2),
         civlib = round(civlib, 2), 
         gni = round(gni, 2))

mean_gini_civ_gni_long <- mean_gini_civ_gni_long %>% 
  mutate(year = as.numeric(year))

plot_3D <- plot_ly(mean_gini_civ_gni_long, 
                   x = ~civlib, 
                   y = ~gini, 
                   z = ~gni, 
                   symbol = ~continent, 
                   symbols = c('circle', 'square', 'diamond', 'cross', 'x', 'triangle-up', 'triangle-down'), 
                   type = 'scatter3d', 
                   mode = 'markers',
                   hovertemplate = paste("CL index: %{x}", "<br>Gini Index: %{y}", "<br>GNI per capita: %{z}"),
                   marker = list(size = 5, 
                                 color = ~year, 
                                 colorscale = 'Viridis', 
                                 showscale = TRUE,
                                 colorbar = list(x = 1.4, y = 0.45, tickmode = "array", tickvals = seq(1975, 2021, 10), ticktext = c("1975", "1985", "1995", "2005", "2015", "2021")))) %>%
  layout(
    title = list(
      text = "3D Scatter Plot: Civil Liberties, Gini Index and GNI per capita",
      y = 0.98),
    scene = list(
      xaxis = list(title = 'Civil Liberties (CL) Index'),
      yaxis = list(title = 'Gini Index'),
      zaxis = list(title = 'Gross National Product per capita')
      ),
    legend = list(
      x = 1.00,
      y = 0.1,
      orientation = "v")
    )
```

```{r reg lines}
smooth_gini_loess_plot <- ggplot(gini_civlib_gni, aes(x = civlib_index, y = gini_index)) +
  geom_point() +
  geom_smooth(method = "loess", col = "blue", se = T) +
  labs(title = "Linear regression: Gini index and Civil Liberty Index",
       x = "Civil Liberty Index",
       y = "Gini Index",
       color = "GNI per capita") +
  theme_minimal()

smooth_gini_lm_plot <- ggplot(gini_civlib_gni, aes(x = civlib_index, y = gini_index)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue", se = T) +
  labs(title = "Linear regression: Gini Index and Civil Liberty Index",
       x = "Civil Liberty Index",
       y = "Gini Index",
       color = "GNI per capita",
       caption = "The shaded area corresponds to the 95% confidence interval.") +
  theme_minimal()

smooth_gini_loess_plotly <- ggplotly(smooth_gini_loess_plot)

smooth_gini_lm_plotly <- ggplotly(smooth_gini_lm_plot)
```

```{r Multivariate analysis attempt}
# trying to find a decent model
# a quadratic equation explains substantially better the variance of the model than with a linear one (gain of approx. 0.05), but a cubic equation seems to already lead to overfitting the data as the gain in variance explained is only  approx. 0.005. However, the variance explained remains low, and it seems more controls should be included in the model in order to better explain the variance. Note that statistical significance is high in all these 3 models, but it shouldn't be surprising given the high amount of data fed into it.

polymodel_data <- na.omit(gini_civlib_gni) 

#model_poly_1 <- lm(gini_index ~ poly(civlib_index, 1) + poly(gni_cap, 1), data = polymodel_data)
model_poly_2 <- lm(gini_index ~ poly(civlib_index, 2) + poly(gni_cap, 2), data = polymodel_data)
#model_poly_3 <- lm(gini_index ~ poly(civlib_index, 3) + poly(gni_cap, 3), data = polymodel_data)
modelsummary <- summary(model_poly_2)

# creating dataframe for predicted data
prediction_data <- data.frame(
  civlib_index = seq(min(polymodel_data$civlib_index), max(polymodel_data$civlib_index), length.out = 100),
  gni_cap = mean(polymodel_data$gni_cap)) # Fixing gni_cap set as control

# inputting predicted data in the prediction dataframe
predictions <- predict(model_poly_2, newdata = prediction_data, se.fit = T)
prediction_data$gini_index <- predictions$fit
prediction_data$se <- predictions$se.fit
prediction_data$lower <- prediction_data$gini_index - 1.96 * prediction_data$se
prediction_data$upper <- prediction_data$gini_index + 1.96 * prediction_data$se

highest_gini <- prediction_data %>% 
  summarise(max_gini = max(gini_index))

polymodel_data <- polymodel_data %>% 
  mutate(tooltip_label = paste("CL Index:", civlib_index,
                               "<br>Gini Index:", gini_index,
                               "<br>GNI per capita:", gni_cap))

# creating regression line
reg_graph <- ggplot(data = polymodel_data, aes(x = civlib_index, y = gini_index)) +
  geom_point(aes(color = gni_cap, text = tooltip_label)) +
  geom_line(data = prediction_data, aes(x = civlib_index, y = gini_index), color = "blue") +
  geom_ribbon(data = prediction_data, aes(x = civlib_index, ymin = lower, ymax = upper), alpha = 0.2) +
  labs(title = "Regression analysis: Gini and CL Indexes, GNI per capita",
       x = "Civil Liberty (CL) Index",
       y = "Gini Index",
       color = "GNI per capita",
       caption = "The shaded area corresponds to the 95% confidence interval.") +
  theme_minimal()

reg_graph_plotly <- ggplotly(reg_graph, tooltip = "text")
```

# SNAPSHOT

## {height=30%}

### {width=40%}
```{r}
# value box with highest and lowest gini in 2023
value_box(
  title = "Top 5 country with the highest Gini Index, 2023",
  value = tags$div(
    style = "font-size: 24px;",
    "Highest Gini Index: South Africa (0.66)",
    tags$br(),
    tags$span(
      style = "font-size: 20px;",  # Adjust the font size as needed
      "Mean Gini Index (worldwide):", mean_gini
    )
  ),
  showcase = sparkcol_highest_gini,
  full_screen = TRUE,
  theme = value_box_theme(bg = "#1E90FF")
)
```


### {width=30%}

```{r}
value_box(
  title = "Top 5 country with the highest Civil Liberties Index, 2023",
  value = tags$div(style = "font-size: 20px;", "Highest Civil Liberties Index: Estonia (91.0)"),
  showcase = sparkcol_highest_civlib,
  full_screen = TRUE,
  theme = value_box_theme(bg = "darkgreen")
)
```

### {width=30%}

```{r}
value_box(
  title = "Top 5 country with the lowest Civil Liberties Index, 2023",
  value = tags$div(style = "font-size: 20px;", "Lowest Civil Liberties Index: North Korea (5.78)"),
  showcase = sparkcol_lowest_civlib,
  full_screen = TRUE,
  theme = value_box_theme(bg = "darkgreen")
)
```


## height=60%}

### {width=50%}
```{r}
map_gini_plotly
```

### {width=50%}
```{r}
map_civlib_plotly
```

# LONGITUDINAL DATA

## {heigth=80%}

### {width=50%}
```{r}
graph_mean_gini_all_plotly
```

### {width=50%}
```{r}
plot_3D
```

## {height=20%}
As shown in the line graph, Gini Indexes of each regions follow different patterns over time. However, it is possible to identify some trends among these regions, with a low rate of economic inequalities between 1975 and 1990, followed more recently by a general increase in economic inequalities since 2018 approximately. 

As shown in the 3D scattered plot, a trend between regions is more difficult to interpret. Indeed, the rate of regions with similar income level range tends to vary quite a lot: for example, Europe have seen its rate of civil liberties increase, while it remained fairly similar in North America. similarly, while civil liberties increased in Africa over time, it actually decreased in Latin America - however, note that the rate of civil liberties remained higher in Latin America than in Africa at any time point. Concerning GNI per capita, trends are more present, with an fairly constant increase of this rate in each regions (although at various speed), followed by a slight decrease in the recent years (which thus seem to be inversely related to Gini Index, as this later indice increases in the recent years).

# MULTIVARIATE ANALYSIS

## {layout-ncol=2}

### {width=50%}

```{r fig.cap="Gini Index and Civil liberties Index are respectively set as dependent and independent variables. The shaded area corresponds to the 95% confidence interval."}
reg_graph_plotly
```


### {width=50%}

#### Description of the model{height=50%}
The model on the left side of the screen displays a non-linear regression model, with GNI per capita used as control. Below is found the summary of the chosen model. Both of these elements shows that while holding GNI per capita constant, an increase in civil liberties lead to an increase in economic inequalities (i.e. higher Gini index), but only to a certain point (`r round(highest_gini, 4)`) after which inequalities decline. 

A quadratic equation has been chosen here because, on the one hand, it explains substantially better the variance (i.e. Multiple R Squared value in the summary of the model below) of the model than with a linear one (gain of approx. 0.05). On the other hand, increasing the complexity with an equation including 3 or more polynomial terms seemingly leads to overfitting, as the gain in variance explained is only  approx. 0.005 in the case of a cubic equation. Despite choosing a quadratic equation appearing to be the best fit for the data, the variance explained remains fairly low (i.e 0.1859), which indicates that only 19% of the total variation of the dependent variable (i.e. Gini index) is explained by the independent variable (i.e. civil liberties), accounting for GNI per capita - this is a relatively low value, and this despite R-squared value related to social sciences related phenomena can be expected to be relatively lower than in other fields such as biomedical sciences. More controls should be included in the model in order to better explain the total variance of the Gini index. 

Note that statistical significance of each coefficient is high (p-value = 0 in each case, as shown in the model's outputs summary section), but this isn't surprising given the fairly high amount of data fed into it.

#### {height=50%}
```{r title="Summary of the regression model"}
modelsummary
```

# DATASET

The data used in this dashboard is shown below, and can be downloaded as a CSV.

```{r}
gini_civlib_gni_clean <- gini_civlib_gni %>% 
  select(country, continent, year, gini_index, civlib_index, gni_cap, - country_code) %>% 
  rename(Country = country,
         Continent = continent,
         Year = year,
         "Gini Index" = gini_index,
         "Civil Liberties Index" = civlib_index,
         "GNI per capita" = gni_cap) %>% 
  mutate(Year = as.character(Year)) %>% 
  filter(Year %in% c("1975", "1985", "2000", "2010", "2021"))
```

```{r title="Dataset"}
gini_civlib_gni_clean_reac <- htmltools::browsable( 
  tagList( 
    reactable(gini_civlib_gni_clean, defaultPageSize = 5, elementId = "gini-table", searchable = TRUE, filterable = TRUE),
    tags$button("Download as CSV", onclick = "Reactable.downloadDataCSV('gini-table')") 
    ) 
  )

gini_civlib_gni_clean_reac
```

# About

All data used in this dashboard is taken from the __Gapminder__ website, which is an independent educational non-profit organization based in Sweden. It aims to combat global misconceptions by providing a fact-based worldview that everyone can understand. Founded in 2005 by Hans Rosling, Anna Rosling Rönnlund, and Ola Rosling, Gapminder uses reliable data to create engaging and easy-to-understand visualizations. Gapminder’s most notable tool, Gapminder World, allows users to explore historical data on various social, economic, and environmental indicators through interactive graphs. 

__Description of the variables:__

__Gini Index:__ 
The Gini Index is a indice measuring the level of inequalities in a country - in other terms, it is a statistical measure enabling to observe how a variable (wealth, income, ...) is distributed within a given population. This is calculated through a Lorenz curve, which sets on a x-y axis the cumulative part of the revenues of a population on the y axis, and the cumulative part of the population on the x axis. In case of perfect equality, the Lorenz curve would actually be a straight line with a positive slope starting at the origin. In case of inequality, the data will plot a curve that sets an area below the line of equality, and the size of this area corresponds to the Gini index. This index varies between 0 and 1, 0 corresponding to perfect equality while 1 corresponds to perfect inequality. Here is a video describing the calculation between this index:

<iframe width="560" height="315" src="https://www.youtube.com/embed/y8y-gaNbe4U?si=of0Dji7C2luME35f" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>
 	
__Civil Liberties Index:__
The civil liberties subattribute denotes the extent to which civil rights and liberties are respected. The five civil liberties sub-components are freedom of expression, freedom of association and assembly, freedom of religion, freedom of movement, and personal integrity and security, each of which reflects core concepts in the human rights literature. The five sub-components were aggregated into the civil liberties subattribute using BFA.

__GNI per capita:__
GNI per capita (formerly GNP per capita) is the gross national income, converted to U.S. dollars using the World Bank Atlas method, divided by the midyear population. GNI is the sum of value added by all resident producers plus any product taxes (less subsidies) not included in the valuation of output plus net receipts of primary income (compensation of employees and property income) from abroad.


